---
title: "Creating an Interactive Map"
subtitle: "Getting Started with tmap"
author: "Conner Surrency, July 2020"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, tamu, tamu-fonts]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r setup, echo=F}
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = FALSE, fig.retina = 3)

library(flipbookr)
```


# What Are We Doing?
--

- We are going to walk through the process of creating a simple interactive map using tmap.
--

- We will start by walking through the steps of creating a simple standard map. After that, we'll transform it into an interactive one.
--

- In this case, we will narrow the data down to only information related to Australia, and then display each city in the nation based on its population. 

---

## Set-Up
--

We will need the following packages (and data from them) installed:  
--

- tidyverse
--

- tmap

---

`r chunk_reveal(chunk_name = "standard", break_type = "user", widths = c(50,50), title="## Building a Simple Map")`

```{r standard,include=F}
library(tidyverse) 
library(tmap) 

## We will be using data from "world" and "Metro"
data(World, metro) 

## We will filter the metro datasets down so that we only have 
## incidences within Australia
metro %>% #BREAK
  filter(iso_a3 == "AUS") -> #BREAK
  metro_australia #BREAK

## Setting the mode as "Plot" will allow us to create a static map
tmap_mode("plot") 

World %>%    #BREAK
## Here we will filter the "World" set down to just Australia as well
  filter(name == "Australia") %>%  #BREAK
## Here we decide the projection type and create polygon shape to 
## represent the country
  tm_shape(shp = .,  
           projection = "+proj=merc") + 
    tm_polygons("name", 
                legend.show = F) +   #BREAK
## Finally we add the datapoints representing cities in Australia
  tm_shape(metro_australia) +
    tm_symbols(col = "orange", 
               size = "pop2020", 
               scale = 1, 
               legend.size.show = T) #BREAK
```

---

## Transforming into an Interactive Map
--

- Turning a static map into an interactive one is extremely easy with tmap.
--

- Simply change the tmap_mode() from "plot" to "view", that's all there is to it.
--

- After we've changed the map mode, you can go ahead and explore the new interactive map. Click-and-drag, scroll in-or-out, toggle layers, and click on data points for more information.

---

`r chunk_reveal(chunk_name = "interactive", break_type = "rotate", widths = c(50,50), title="## Transforming to Interactive")`

```{r interactive,include=F}
library(tidyverse) 
library(tmap) 

data(World, metro) 

metro_australia <- metro %>% filter(iso_a3 == "AUS") 

## Swapping "plot" for "view" transforms the map from static 
## to interactive
tmap_mode("plot") #ROTATE
tmap_mode("view") #ROTATE

## Simple as that. Try exploring the map now, drag, scroll, click

World %>%   
  filter(name == "Australia") %>%  
  tm_shape(shp = .,  
           projection = "+proj=merc") + 
    tm_polygons("name",  
                legend.show = F) +   
  tm_shape(metro_australia) +
    tm_symbols(col = "orange", 
               size = "pop2020", 
               scale = 1, 
               legend.size.show = T) 
```
---
## Bonus Content: Map Projections
--

- By default, the projection type will be defined by the projection used by the shp object itself. However, its easy to manually switch out projections in tmap as an argument of tm_shape.
--

- We will rotate through a few good projection types on our first static map.

---

`r chunk_reveal(chunk_name = "worldmap", break_type = "rotate", widths = c(50,50), title="## Projections")`

```{r worldmap, eval = F, echo = F}
library(tidyverse) 
library(tmap) 

data(World, rivers) 

tmap_mode("plot") 

World %>%   
  tm_shape(shp = .,  
## Mercator
           projection = "+proj=merc") + #ROTATE 
## Robinson
           projection = "+proj=robin") +  #ROTATE 
## Wagner I
           projection = "+proj=wag1") +  #ROTATE 
## Interrupted Goode Homolosine
           projection = "+proj=igh") +    #ROTATE 
## Nicolosi Globular
           projection = "+proj=nicol") +   #ROTATE 
## Lagrange
           projection = "+proj=lagrng") +  #ROTATE
## Van Der Grinten I
           projection = "+proj=vandg") +    #ROTATE 
## Tobler-Mercator
           projection = "+proj=tobmerc") +  #ROTATE 
## American Polyconic
           projection = "+proj=poly") +   #ROTATE 
## Orthographic
           projection = "+proj=ortho") +  #ROTATE 
## Lambert Equal Area Conic
           projection = "+proj=leac") +     #ROTATE 
    tm_polygons("name",  
                legend.show = F) +   
  tm_shape(rivers) +
    tm_lines("lightblue", 
             lwd = "strokelwd", 
             scale = 3, 
             legend.lwd.show = F) 

```



---

## Additional Information & Options
--

- Different projections are better suited to different maps. 
--

- For example, if producing a map of Alaska then the "Alaska Grid" or "Alaska Series E" projection type would be typical. 
--

- When presenting a large region or the world at large, however, "Mercator" or "Robinson" are more reliable. 
--

- Find more projection types at: https://proj.org/operations/projections/index.html


