---
title: "Introduction to gganimate"
subtitle: "Unemployment in Europe"
author: "Luca Picci"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---




```{r, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scratch can be time consuming
#knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = TRUE, fig.retina = 3)
library(flipbookr)
library(tidyverse)
library(gganimate)
library(ggplot2)
library(readxl)
```


```{r, Unempplot, include=FALSE}
#2005 - 2017
unemployment <- read_excel("C:\\Users\\lpicc\\OneDrive\\Documents\\python work\\unemp.xlsx")

```


# Getting Started
Showing the steps to create an animated bar chart can be tricky.
In order to demonstrate the steps required to build this annimated plot we will:
- Build the skeleton of the plot (declare aesthetics and build the theme and annotations)
- Then use facet_wrap() and declare the geoms to show the different bar lengths at each time period
- Finally we will combine the plots and add the animation


---
`r chunk_reveal("plotdata", title = "Data Manipulation")`

```{r, plotdata, include=FALSE}
unemployment %>% 
  filter(SUBJECT == "TOT") %>% 
  filter(TIME>=2007) %>% 
  filter(Country!= "Estonia")%>% 
  select(-LOCATION) %>% 
  select(-SUBJECT) ->
  plotdata
```


---
`r chunk_reveal("plot1", title = "Building the Plot")`

```{r, plot1, include=FALSE}
ggplot(data = plotdata) +
  aes(y = reorder(Country, Value))+
  aes(x = Value)+
  geom_segment(aes(x = 0, xend = 65, 
                   yend = Country),
               size = 7.5, 
               color = "azure3")+
  scale_x_continuous(position = "top", 
                     breaks = c(0, 25, 50), 
                     labels=
                       function(x) paste0(x,"%"))+
  labs(title = "Euro Area Youth Unemployment Rate 2007 - 2017")+
  labs(subtitle = "Youth unemployment (15-24 years) expressed as a percent of the youth labour force")+
  labs(y = NULL)+
  labs(caption = "*excl. Malta, Cyprus, Estonia\nSource:OECD (2020), Youth unemployment rate (indicator)")->
  plot
 
```




---

`r chunk_reveal("plot2", title = "Building the Theme")`

```{r, plot2, include=FALSE}
plot+
  theme(panel.background = 
      element_rect(fill = "grey99"))+
  theme(plot.background = 
      element_rect(fill = "grey99"))+
  theme(panel.grid = element_blank())+
  theme(plot.title = 
      element_text(size = 18, hjust = 0, 
                   color = "steelblue4", 
                   face = "bold"))+
  theme(plot.subtitle = 
      element_text(size = 12, 
                   color = "darkgoldenrod2", 
                   hjust = 0, face = "bold"))+
  theme(plot.caption = 
      element_text(color = "steelblue4", 
                   hjust = 0.94))+
  theme(axis.title.x = 
      element_text(size = 12, hjust = 0, 
                   color = "steelblue4", 
                   face = "bold"))+
  theme(axis.text.x = 
      element_text(size = 10, 
                   color = "steelblue4", 
                   hjust = 0.5))+
  theme(
    axis.line.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y= element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank())->
  plot2
```

---
# Faceting 
This intermediate step is not necessary, but in order to demostrate clearly how the plot will be built, we will facet by year using facet_wrap() to create an individual plot for each year.
This will allow us to see how the bar lengths vary and add annotations before adding animations
---


`r chunk_reveal("plot3", title = "Faceting")`


```{r, plot3, include=FALSE}
plot2+
  facet_wrap(~ TIME, nrow=3)+
  geom_col(fill = "steelblue4", 
           width = 0.75, 
           alpha = 0.9)+
  annotate(x=0, xend=65, 
           y=15, yend=15, 
           colour="steelblue4", 
           geom="segment", 
           size = 2)+
  annotate(x = 25, xend = 25, 
           y = 14.9, yend = 0.5, 
           size = 1, geom="segment", 
           color = "darkgoldenrod2", 
           linetype = "dashed")+
  annotate(x = 50, xend = 50, 
           y = 14.9, yend = 0.5, 
           size = 1, geom="segment", 
           color = "darkgoldenrod2", 
           linetype = "dashed")->
  plot3
```





---
`r chunk_reveal("plot4", break_type = "user", title="Adding the Animation", widths = c(22, 25), display_type = c("code", "output"))`


```{r, plot4, include=FALSE}
plot3+ #BREAK
  facet_null()+
  gganimate::transition_time(TIME)+ #BREAK
  ease_aes('linear')+#BREAK
  labs(x = 'Year: {format(round(frame_time, digits = 0))}')+
  geom_text(aes(label = reorder(Country, Value)), 
            x = 0.1, hjust = 0, color = "grey99", 
            fontface = "bold")+ #BREAK
  geom_text(aes(label = 
                  format(round(Value, digits = 0))),
            x = 63, color = "steelblue4", 
            fontface = "bold", hjust = 1)+ #BREAK
  geom_text(x=63.5, label = "%", color = "steelblue4", 
            fontface = "bold", hjust = 0)-> #BREAK
  finalplot

```

---
`r chunk_reveal("finplot", display_type = "output")`

```{r, finplot, include=FALSE}
finalplot

```





```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.2; font-size: 73%}

@media print {
  .has-continuation {
    display: block;
  }
}
```
